paper_replication

### 2021-12-14 paper_replication.Rmd
Update paper_replication.Rmd

TODO: model the propensity weights
    - If I understand this right, we're modeling the propensity for a county to be located in a highly restrictive state.  Is that right?  So we want an estimated probability of a binary outcome, so we want logistic regression. Is that right?

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#load packages
options(tidyverse.quiet=TRUE)
library(tidyverse)
library(readxl)
library(dplyr)
library(geosphere)
library(stringr)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#load in data
providers <- read.csv("provider_list_geo.csv")
countycentroids <- read.csv("county_pop_centroids.csv")
countydem <- read.csv("county_demographics.csv")
abortions <- read.csv("cleanedabortiondata.csv")
restrictions <- read.csv("state_restrictions2010.csv")
countypres <- read.csv("countypres.csv")
```

```{r}
########### calculate abortion rates in 2010 ###########
abortions2010 <- abortions %>% dplyr::select(state, county, X2010)

# clean countydem data to make counties and states match abortion count data 
countydem <- read.csv("county_demographics.csv")

@@

abortions2010$county <- gsub(abortions2010$county, pattern = "[()]", replacement = "")
abortions2010$county <- gsub(abortions2010$county, pattern = "ind. city", replacement = "city")
abortions2010$county <- gsub(abortions2010$county, pattern = "Saint", replacement = "St.")
abortions2010$county <- gsub(abortions2010$county, pattern = "Dekalb", replacement = "DeKalb")
abortions2010$county <- gsub(abortions2010$county, pattern = "Desoto", replacement = "DeSoto")
abortions2010$county <- gsub(abortions2010$county, pattern = "Miami Dade", replacement = "Miami-Dade")
abortions2010$county <- gsub(abortions2010$county, pattern = "Hawai'i", replacement = "Hawaii")
abortions2010$county <- gsub(abortions2010$county, pattern = "Kaua'i", replacement = "Kauai")
abortions2010$county <- gsub(abortions2010$county, pattern = "City", replacement = "city")
countydem$county <- gsub(countydem$county, pattern = "City", replacement = "city")
abortions2010$county <- gsub(abortions2010$county, pattern = "Lac Qui Parle", replacement = "Lac qui Parle")
abortions2010$county <- gsub(abortions2010$county, pattern = "St.e Genevieve", replacement = "Ste. Genevieve")
abortions2010$county <- gsub(abortions2010$county, pattern = "Shannon/Oglala Lakota", replacement = "Oglala Lakota")
abortions2010$county <- gsub(abortions2010$county, pattern = "Lamoure", replacement = "LaMoure")

for (i in 1:nrow(abortions2010)) {
  if (length(countydem$Women[countydem$state.abb == abortions2010$state[i] & countydem$county == abortions2010$county[i]]) == 0) {
    abortions2010$Women[i] <- NA
  }
  else {
    abortions2010$Women[i] <- countydem$Women[countydem$state.abb == abortions2010$state[i] & countydem$county == abortions2010$county[i]]
  }
}

@@

abortions2010 <- abortions2010 %>% filter(!(county %in% c("NONRESIDENTS", "PROTECTED", "OTHER STATES", "OTHER COUNTRIES", "NON-RESIDENTS", "OTHER", "NEVADA PROTECTED", "NEVADA UNSPECIFIED", "NEBRASKA PROTECTED", "county", "MARYLAND", "St. Louis County")))

abortions2010missingwomen <- abortions2010[is.na(abortions2010$Women), ]

#make rates
abortions2010 <- abortions2010 %>% mutate(rate = X2010/Women)

@@

```{r}
## calculate percent democratic votes
countypresdem2008 <- countypres %>% filter(party == "DEMOCRAT" & year == 2008)
countypresdem2008$county_name <- str_to_title(countypresdem2008$county_name)
countypresdem2008$county_name[!(countypresdem2008$county_name %in% countydem$county)]

countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = "City", replacement = "city")
countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = "Dekalb", replacement = "DeKalb")
countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = "Desoto", replacement = "DeSoto")
countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = "Lac Qui Parle", replacement = "Lac qui Parle")

countypresdem2008 <- countypresdem2008 %>% mutate(percentdem2008 = candidatevotes/totalvotes) 
names(countypresdem2008)
countypresdem2008 <- countypresdem2008 %>% rename(state.abb = state_po, county = county_name) %>% dplyr::select(state.abb, county, percentdem2008)

@@

```{r}
# calculate distances from county centroids to closest provider location

@@

```{r}
#make propensity scores

>>>>

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#load packages
options(tidyverse.quiet=TRUE)
library(tidyverse)
library(readxl)
library(dplyr)
library(geosphere)
library(stringr)
library(kableExtra)
library(table1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#load in relevant data
providers <- read.csv("provider_list_geo.csv")
countycentroids <- read.csv("county_pop_centroids.csv")
countydem <- read.csv("county_demographics.csv")
abortions <- read.csv("cleanedabortiondata.csv")
restrictions <- read.csv("state_restrictions2010.csv")
countypres <- read.csv("countypres.csv")
```

```{r}
########### calculate abortion rates in 2010 ###########

#filter abortion rate data to only 2010
abortions2010 <- abortions %>% dplyr::select(state, county, X2010)

# clean countydem data to make counties and states match those in abortion count data 
countydem <- read.csv("county_demographics.csv")

@@

abortions2010$county <- str_to_title(abortions2010$county)
countydem$county <- str_to_title(countydem$county)

abortions2010$county <- gsub(abortions2010$county, pattern = "[()]", replacement = "")
abortions2010$county <- gsub(abortions2010$county, pattern = "Ind. City", replacement = "City")
abortions2010$county <- gsub(abortions2010$county, pattern = "Saint", replacement = "St.")
abortions2010$county <- gsub(abortions2010$county, pattern = "Miami Dade", replacement = "Miami-Dade")
abortions2010$county <- gsub(abortions2010$county, pattern = "Hawai'i", replacement = "Hawaii")
abortions2010$county <- gsub(abortions2010$county, pattern = "Kaua'i", replacement = "Kauai")
abortions2010$county <- gsub(abortions2010$county, pattern = "St.e Genevieve", replacement = "Ste. Genevieve")
abortions2010$county <- gsub(abortions2010$county, pattern = "Shannon/Oglala Lakota", replacement = "Oglala Lakota")

#Add number of women to abortion count data
for (i in 1:nrow(abortions2010)) {
  if (length(countydem$Women[countydem$state.abb == abortions2010$state[i] & countydem$county == abortions2010$county[i]]) == 0) {
    abortions2010$Women[i] <- NA
  }
  else {
    abortions2010$Women[i] <- countydem$Women[countydem$state.abb == abortions2010$state[i] & countydem$county == abortions2010$county[i]]
  }
}

@@

#remove non-counties
abortions2010 <- abortions2010 %>% filter(!(county %in% c("Nonresidents", "Protected", "Other States", "Other Countries", "Non-Residents", "Other", "Nevada Protected", "Nevada Unspecified", "Nebraska Protected", "County", "Maryland")))

#look at counties missing women pop data
abortions2010missingwomen <- abortions2010[is.na(abortions2010$Women), ]
#DeWitt, Bedford city, Clifton Forge city in abortion rate data  but not county dem data

#calculate abortion rates
abortions2010 <- abortions2010 %>% mutate(rate = X2010/Women)

#Add abortion counts to county data
for (i in 1:nrow(countydem)) {
  if (length(abortions2010$rate[abortions2010$state == countydem$state.abb[i] & abortions2010$county == countydem$county[i]]) == 0) {
    countydem$AbortionRate2010[i] <- NA
  }
  else {
    countydem$AbortionRate2010[i] <- abortions2010$rate[abortions2010$state == countydem$state.abb[i] & abortions2010$county == countydem$county[i]]
  }
}

@@

```{r}
########### calculate percent democratic votes by county in in 2008 ###########
countypresdem2008 <- countypres %>% filter(party == "DEMOCRAT" & year == 2008)
countypresdem2008$county_name <- str_to_title(countypresdem2008$county_name)

countydem$county <- gsub(countydem$county, pattern = " Parish", replacement = "")
countydem$county <- gsub(countydem$county, pattern = " City", replacement = "")
countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = " City", replacement = "")
countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = "Saint", replacement = "St.")
countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = " County", replacement = "")

missingcounties <- countypresdem2008[!(countypresdem2008$county_name %in% countydem$county), ]

countypresdem2008 <- countypresdem2008 %>% mutate(percentdem2008 = candidatevotes/totalvotes) 
countypresdem2008 <- countypresdem2008 %>% rename(state.abb = state_po, county = county_name) %>% dplyr::select(state.abb, county, percentdem2008)

@@

```{r}
########### calculate distances from county centroids to closest provider location ###########

@@

```{r}
#look at demographics

countydem <- countydem %>% mutate(inanalysis = case_when(state.abb %in% c("AZ", "DE", "GA", "IL", "IN", "MI", "NY", "NC", "OH", "OK", "OR", "PA", "SC", "TX", "UT", "VT", "WA", "WI") ~ 1, !(state.abb %in% c("AZ", "DE", "GA", "IL", "IN", "MI", "NY", "NC", "OH", "OK", "OR", "PA", "SC", "TX", "UT", "VT", "WA", "WI")) ~ 0))

countydemanalysis <- countydem %>% filter(inanalysis == 1)

summarytable <- countydemanalysis %>% group_by(state) %>% summarize(totalcounties = n())
usedcounties <- aggregate(AbortionRate2010 ~ state, data = countydemanalysis, function(x) {sum(!(is.na(x)))}, na.action = NULL)
summarytable <- left_join(summarytable, usedcounties, by = "state")
summarytable %>% mutate(percent = AbortionRate2010/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Abortion Data in 2010", "Percent"), caption = "Table 1. Number of Counties with Abortion Data Included in Analyses by State") %>% kable_styling()
names(countydem)
countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)

countydem$inanalysislabel <- factor(countydem$inanalysis, levels = c(1, 0), labels = c("Counties in States With Abortion Data", "Counties in States Without Abortion Data"))

label(countydem$Women) = "Women"
label(countydem$White.pct) = "White"
label(countydem$Black.pct) = "Black"
label(countydem$Native.American.pct) = "Native American"
label(countydem$Asian.pct) = "Asian"
label(countydem$Hispanic.pct) = "Hispanic"
label(countydem$Median.Income) = "Median Income"
label(countydem$College.Grad.pct) = "College Graduates"
label(countydem$percentdem2008) = "Voting for the Democratic candidate in the most recent presidential election"

units(countydem$Women) = "No."
units(countydem$White.pct) = "%"
units(countydem$Black.pct) = "%"
units(countydem$Native.American.pct) = "%"
units(countydem$Asian.pct) = "%"
units(countydem$Hispanic.pct) = "%"
units(countydem$Median.Income) = "$"
units(countydem$College.Grad.pct) = "%"
units(countydem$percentdem2008) = "%"

table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | inanalysislabel, data = countydem)

propensitywts <- glm(inanalysis ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008, data=countydem, family=binomial())
summary(propensitywts)

p.score <- predict(propensitywts, countydem, type = "response")
df <- countydem
df$ps <- p.score

#3 plots of distribution of propensity scores

ggplot(df, aes(x = ps, group = inanalysislabel)) + geom_density(aes(fill = inanalysislabel), alpha = 0.2) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical') +
  ylab("Density")

ggplot(df, aes(x = ps, fill = inanalysislabel, color = inanalysislabel)) +
  geom_histogram(alpha = 0.3, position = 'identity', bins=15) +
  facet_grid(as.factor(inanalysis) ~ .) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  scale_color_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical')

df %>%
  mutate(ps.grp = round(ps/0.05) * 0.05) %>%
  group_by(inanalysis, ps.grp) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(n2 = ifelse(inanalysis == 0, yes = n, no =  -1*n)) %>%
  ggplot(aes(x = ps.grp, y = n2, fill = as.factor(inanalysis))) +
  geom_bar(stat = 'identity', position = 'identity') +
  geom_text(aes(label = n, x = ps.grp, y = n2 + ifelse(inanalysis == 0, 8, -8))) +
  xlab('Probability of Being in the Analysis') +
  ylab('N') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete(name = "", labels = c("Counties in States With Abortion Data", "Counties in States Without Abortion Data")) +
  scale_x_continuous(breaks = seq(0, 1, 0.05)) +
  theme(legend.position = 'bottom', legend.direction = 'vertical',
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank())

```{r}
countydemanalysis <- df %>% filter(inanalysis == 1)
countydemanalysis <- left_join(countydemanalysis, restrictions, by = "state")
countycentroids <- countycentroids %>% rename(county = COUNAME, state = STNAME)
countycentroids$county <- str_to_title(countycentroids$county)
countycentroids <- countycentroids %>% select(county, state, distancetoprovider)
countydemanalysis <- left_join(countydemanalysis, countycentroids, by = c("state", "county"))

countydemanalysis$highorlow <- factor(countydemanalysis$highorlow, levels = c("low", "high"))
model <- lm(AbortionRate2010 ~ highorlow + distancetoprovider + highorlow*distancetoprovider, data = countydemanalysis, weights = ps)
summary(model)
